A generic flow system can be described as a system whose state depends on
flowing streams of energy, material or information. Traditional examples of
such systems are the steady transport of properties such as pressure,
temperature or matter in fluids, and the transport of charge in electrical
currents. However, valuable insight into various phenomena, including, but not
limited to, the development of algae in the ocean and crowd patterns formed
by humans, can be obtained by regarding them as flow systems.

Lagrangian coherent structures (henceforth abbreviated to LCSs) can be described
as `landscapes' within a multidimensional space, which exert a major influence
upon the flow patterns in dynamical systems. Such structures provide a
simplified means of predicting the future states of complex systems, like
nonlinear manybody problems, over a time interval of interest, compared to the
conventional approach of increasing the spatial and temporal resolution of the
model(s) involved in numerical simulations. From their variational theory,
hyperbolic LCSs are identified as locally most repelling or attractive material
surfaces. Somewhat simplified, such surfaces can be considered as a
generalization of trajectories created by the underlying transport phenomenon.
Hyperbolic LCSs form the skeleton of observable flow patterns.

This project is centered around the dependence of the numerical identification
of such LCSs upon which numerical integration scheme was used in calculating
them. LCSs were computed and compared for an analytical double gyre velocity
field, for which reference LCSs are documented in the literature. In particular,
four traditional singlestep and four embedded, adaptive stepsize Runge-Kutta
methods were investigated. A numerical solution obtained by means of a high
order embedded method with a very small numerical tolerance --- two orders of
magnitude smaller than the smallest tolerance level utilized otherwise ---
and thus a very small expected numerical error, was utilized as the basis
for comparing the performance of the different methods.
%LCSs were computed and
%compared for an analytical velocity field, for which reference LCSs are
%documented in the literature.

It turns out that the LCSs for the system under consideration are quite robust
numerically, at least for the chosen parameter values. Thus, none of the
numerical integration schemes stand out as the absolutely best-suited. However,
based upon the errors of the computed flow maps, high order integration
methods are generally advisable for generic flow systems, as they invariably
yield more efficient calculations, which are less susceptible to numerical
round-off error. Moreover, there is no apparent reason to expect the LCSs of
a generic flow system to be as robust as they were for the system considered
for this project work.

Note that the numerical implementation of one of the LCS existence conditions
which arise from their variational theory has not yet (to the author's
knowledge) received a complete description in the literature. This is a
consequence of the study of LCSs being in it relative infancy, as a branch of
physics. The approach considered in this project is based on careful inspection
of properties of the underlying system, which may or may not be applicable to
generic systems, depending on their scale. In other words, there is currently a
lot of room for research within the field.
