\newpage
\section{Estimation of errors}
\label{sec:estimation_of_errors}

Aside from the qualitiative visual comparison between the obtained LCSs and
the reference LCS, a set of numerical errors was also computed, with a view to
use them in order to explain any visual discrepancies. To this end, the error
in the flow map, given by equation~\eqref{eq:flowmap}, was computed as the
root mean square deviation (hereafter abbreviated to RMSD) with regards to the
reference, as follows:
\begin{equation}
    \label{eq:rmsdflowmap}
    \rmsd_{\textnormal{flow map}} = \sqrt{\frac{1}{\widetilde{N}}%
        \sum\limits_{\vct{x}_{0}\,\in\,\widetilde{\mathcal{U}}}%
\big(\widehat{\vct{F}}{}_{t_{0}}^{t}(\vct{x}_{0})-\vct{F}_{t_{0}}^{t}{(\vct{x}_{0})}\big)^{2}},
\end{equation}
where the summation is over all of the tracers in the extended computational domain
$\widetilde{\mathcal{U}}$, that is, both the main and auxiliary tracers which
originate from points in the extended grid, as outlined in
\cref{sub:generating_a_set_of_initial_conditions}. $\widetilde{N}$ is the
total number of advected tracers, and
$\widehat{\vct{F}}{}_{t_{0}}^{t}(\vct{x}_{0})$ is the flow map approximated by
the numerical integrator in question.

The RMSD in both sets of eigen\emph{values}, $\lambda_{1}(\vct{x}_{0})$ and
$\lambda_{2}(\vct{x}_{0})$ of the Cauchy-Green strain tensor field, as given
by equation \eqref{eq:cauchygreencharacteristics} was also computed completely
analogously:
\begin{equation}
    \label{eq:rmsdlmbd}
    \rmsd_{\textnormal{eigenvalue}} = \sqrt{\frac{1}{N}%
    \sum\limits_{\vct{x}_{0}\,\in\,\mathcal{U}}%
\big(\widehat{\lambda}(\vct{x}_{0})-\lambda(\vct{x}_{0})\big)^{2}},
\end{equation}
where the summation is over all main tracers in the principal grid, that is,
all main tracers which originate from within (or the borders of) the
computational domain $\mathcal{U}$. $N$ is the total number of main tracers
in the principal grid. $\widehat{\lambda}(\vct{x}_{0})$ is the approximation of
the true eigenvalues $\lambda(\vct{x}_{0})$ located at $\vct{x}_{0}$, obtained
by means of the numerical integrator in question.

The best way of estimating the error in the computed eigen\emph{vectors} of the
Cauchy-Green Strain tensor field is by means of their orientation, because
the eigenvectors are normalized to unit length. Thus, the eigenvector directions
were computed as azimuthal angles. The RMSD for the direction of
the eigenvector fields was then calculated as follows:
\begin{equation}
    \label{eq:rmsddirection}
    \begin{gathered}
        \vct{\xi}(\vct{x}_{0}) = \begin{pmatrix}\xi_{x}(\vct{x}_{0}),&%
            \xi_{y}(\vct{x}_{0})
        \end{pmatrix},\quad\phi(\vct{x}_{0}) = \arctan{\bigg(\frac{\xi_{y}(\vct{x}_{0})}{\xi_{x}(\vct{x}_{0})}\bigg)},\\
        \rmsd_{\textnormal{eigenvector direction}} = \sqrt{\frac{1}{N}%
    \sum\limits_{\vct{x}_{0}\,\in\,\mathcal{U}}%
{\big(\widehat{\phi}{(\vct{x}_{0})}-\phi{(\vct{x}_{0})}\big)}^{2}},
    \end{gathered}
\end{equation}
where the conventions for $\mathcal{U}$ and $N$ are the same as for the
RMSD of the eigenvalues, given in equation~\eqref{eq:rmsdlmbd}.
$\widehat{\phi}(\vct{x}_{0})$ denotes the azimuthal angle of the eigenvector
located at $\vct{x}_{0}$, found by means of
the numerical integrator in question.
%\clearpage

Regarding the error in the computed LCS curves, false positives, that is,
any parts which of the computed LCS curves which are not present in
the reference curve, and false negatives, i.e., any parts of the reference curve
which is not present in the computed LCS curves, should be treated separately.
This is because as they would influence the predictions of overall flow patterns
in the system in different ways. Numerical noise could result in LCS curve
segments being erroneously identified as either a false positive or a false
negative. In order to reduce this sort of error, a lower numerical threshold
$l_{\textnormal{noise}}=0.01$, that is, ten times the numerical integration
step $\Delta=10^{-3}$ used in order to compute the constituent strainlines, was
used in identifying false positives and negatives; any point on an LCS curve
which was farther away from all points on the reference LCS than
$l_{\textnormal{noise}}$ was identified as a false positive.
Similarly, any point on the reference LCS curve which was farther away from all
points on the LCS curve under consideration than $l_{\textnormal{noise}}$ was
flagged as a false negative.

In order to estimate the offset of the false LCS
segments, the area between these segments and the reference LCS curve
was approximated by the midpoint rule for numerical integration;
\begin{equation}
    \label{eq:midpointfalselcs}
    \textnormal{Offset}_{\textnormal{false LCSs}} = %
    \sum\limits_{\substack{\vct{x}\,\in\,\gamma\\\vct{x}\,\notin\,\widehat{\gamma}}}%
\min\limits_{\widehat{\vct{x}}\,\in\,\widehat{\gamma}}%
\norm{\widehat{\vct{x}}-\vct{x}}\cdot\Delta,
\end{equation}
where the sum is over all points with a greater offset than the aforementioned
$l_{\textnormal{noise}}$. Per equation~\eqref{eq:strainlineode}, $\Delta$, the
previously mentioned numerical integration step length used in order to compute
strainlines, equals the distance separating consequent points in the
parametrization of the LCS curves, due to the eigenvectors $\vct{\xi}_{1}$ being
normalized to unit length. The calculation of this offset is illustrated in
figure~\ref{fig:fp_fn_principle}. Concerning false positives, $\widehat{\gamma}$
denotes the reference LCS curve, and $\gamma$ the approximation under
consideration. For false negatives, the roles are reversed.

\input{mainmatter/method/figures/fp_fn_offset.tex}
\clearpage
Lastly, in order to obtain a quantitative measure of the offset in the
LCS curve segments which comply with the reference LCS (that is, to the
numerical threshold $l_{\textnormal{noise}}$), the RMS distance separating each
point on the LCS curve with the nearest point on the reference, was calculated:
\begin{equation}
    \label{eq:rmsdlcs}
    \rmsd_{\textnormal{LCS}} = \sqrt{%
        \frac{1}{\breve{N}}\sum\limits_{\vct{x}\,\in\,\gamma}%
    \min\limits_{\widehat{\vct{x}}\,\in\,\widehat{\gamma}}{\norm{\widehat{\vct{x}}-\vct{x}}}^{2}},
\end{equation}
where the term $\rmsd$ is used somewhat loosely, in order to conform with
notation used for the other measures of error introduced in this section.
$\breve{N}$ is the number of points which constitute
$\widehat{\gamma}$, the LCS curve found by means of the numerical
integrator in question, which are within the distance $l_{\textnormal{noise}}$
of any point on $\gamma$, the reference LCS curve. The idea
is that LCS curves need not necessarily start from nor end at the same points
in space. Thus, a robust way of estimating the error
of a given LCS curve is the summation of the smallest distances
separating each point on it from the reference.
