\subsection{Computed deviations in the strain eigenvalues and -vectors}
\label{sub:computed_deviations_in_the_strain_eigenvalues_and_vectors}

For brevity, only the $\rmsd$ of $\lambda_{2}$ and the direction of
$\vct{\xi}_{2}$, are included here. The latter is justified because of
the orthogonality of the eigenvectors of the Cauchy-Green strain tensor, which
means that the $\rmsd$ of the directions of both eigenvectors \emph{must} be
identical. Furthermore, the $\rmsd$ of $\lambda_{1}$ scales similarly to
that of $\lambda_{2}$, both as a function of numerical step length (for the
singlestep methods) and the number of function evaluations. Additionally, the
numerical reformulation of the existence theorem for hyperbolic LCSs, given in
\cref{eq:numericalexistence}, exhibits a more sensitive dependence to
$\lambda_{2}$ than $\lambda_{1}$. This is explicitly observable from
conditions~\eqref{eq:numericalexistence2} and~\eqref{eq:numericalexistence4},
and favors $\lambda_{2}$ as the most crucial eigenvalue.

Figure~\ref{fig:lmbd2_err_fixed} indicates that the $\rmsd$ of $\lambda_{2}$,
as defined in~\cref{eq:rmsdlmbd}, follows the anticipated scalings with the
step lengths for the various singlestep integrators. Interestingly, the $\rmsd$
dependence on the numerical step length is completely analogous to that of the
$\rmsd$ of the flow maps, as presented in
\cref{sub:computed_deviations_in_the_flow_maps}. Furthermore, inspection of
figure~\ref{fig:lmbd2_err_both} reveals that the $\rmsd$ of the $\lambda_{2}$
scales similarly to the $\rmsd$ of the flow maps for all of the integration
schemes, differing only by a numerical prefactor. This makes sense, seeing as
the calculation of the Cauchy-Green strain tensor field is based upon
finite differencing applied to the flow map, as described in
\cref{sec:calculating_the_cauchy_green_strain_tensor}.

\input{mainmatter/results/figures/error_figures/lmbd2_error_fixed_steplength.tex}

\input{mainmatter/results/figures/error_figures/lmbd2_error_both.tex}

\vspace{\fill}
\newpage

\Cref{fig:xi2_err_fixed} reveals that the $\rmsd$ of the eigenvector direction,
as defined in~\cref{eq:rmsddirection}, follows the anticipated scalings with
the step lengths for the various singlestep integrators to some degree. The
$\rmsd$ dependence on the numerical step length is not entirely analogous to
that of the $\rmsd$ of $\lambda_{2}$. Notably, the error in the eigenvector
directions is two to three orders of magnitude smaller than the error
in $\lambda_{2}$, for any given step size. This can be understood as a
consequence of using the auxiliary tracers in order to compute the strain
eigenvectors, as described in
\cref{sec:calculating_the_cauchy_green_strain_tensor}, which, by construction,
is more accurate than using the main tracers (which are used to
compute the strain eigenvalues). \Cref{fig:xi2_err_both} shows that the $\rmsd$ of
the eigenvector direction behaves qualitatively like that of the
$\rmsd$ of $\lambda_{2}$ (see~\cref{fig:lmbd2_err_both}) as a function
of the required number of function evaluations for each integration scheme.
Like for the singlestep methods, the $\rmsd$ of the eigenvector direction for
the embedded methods is generally between two and three orders of magnitude
smaller than the $\rmsd$ of $\lambda_{2}$ for any given number of function
evaluations.

%Figure~\ref{fig:xi2_err_fixed} reveals that the $\rmsd$ of the eigenvector
%direction, as defined in~\cref{eq:rmsddirection}, is negligible for
%all numerical step lengths. In fact, the numerical errors are comparable
%to, or even smaller than, the machine epsilon of double-precision floating-point
%numbers, which, as mentioned in
%\cref{sub:generating_a_set_of_initial_conditions}, is of order $10^{-16}$
%\parencite{ieee2008standard}. Because these errors are so small, the agreement
%with the anticipated scalings with the stepsize should not necessarily be taken
%as conclusive evidence of general behaviour. Figure~\ref{fig:xi2_err_both},
%shows that this is in fact the case for the adaptive stepsize methods too,
%aside from the tolerance level $\textnormal{tol}=10^{-1}$, which corresponds to
%the fewest function evaluations for each scheme. As has already been
%established in figure \ref{fig:u0_dom_errs}, for that tolerance level, none of
%the integrators yield $\mathcal{U}_{0}$ domains with reasonable degrees of
%resemblance to the reference, which is shown in figure~\ref{fig:u0_domain}.
%
From \cref{fig:lmbd2_err_both,fig:xi2_err_both}, it is clear that the
$\rmsd$s of $\lambda_{2}$ and the eigenvector direction are both quite large
for the tolerance level $\textnormal{tol}=10^{-1}$ (corresponding to the
smallest number of function evaluations) for all of the embedded methods.
This seems reasonable, given the erroneous $\mathcal{U}_{0}$ domains
obtained for that tolerance level, shown in~\cref{fig:u0_dom_errs} (compare
to~\cref{fig:u0_domain}). Moreover, the sharp turns which are present in
some of the computed LCS curves, perhaps most prominently in
\cref{fig:lcs_rkbs54,fig:lcs_rkdp54} for the tolerance level
$\textnormal{tol}=10^{-2}$, are likely a consequence of the error in the
eigenvalues and -vectors being sufficient for the corresponding
$\mathcal{U}_{0}$ domains to extend to regions in which there are very
strong local orientational discontinuities in the $\vct{\xi}_{1}$-field.
There, the special-purpose linear interpolation introduced in
\cref{sub:a_framework_for_computing_smooth_strainlines} is evidently
insufficient for computing smooth strainlines.

%Aside from the least accurate tolerance level for the embedded integrators,
%though, the error in eigenvector direction is comparable to, or smaller than,
%double-precision machine epsilon. Thus, we may conclude that the eigenvector
%directions are computed correctly for just about any tolerance level and
%numerical time step length. This can be understood as a consequence of the use
%of the auxiliary tracers in order to compute the strain eigenvectors, as
%described in~\cref{sec:calculating_the_cauchy_green_strain_tensor}, which, by
%construction, is more accurate than using the main tracers. Moreover, these
%results imply that the error in the resulting LCS curves are strongly driven by
%the error in the computed eigenvalues. The sharp turns in the computed LCS
%curves for the tolerance level $10^{-2}$, which are particularly visible
%in~\cref{fig:lcs_rkbs54,fig:lcs_rkdp54}, are likely a consequence of the
%corresponding $\mathcal{U}_{0}$ domains extending to regions in which
%there are very strong local orientational discontinuities in the
%$\vct{\xi}_{1}$-field, for which the special-purpose linear interpolation
%introduced in~\cref{sub:a_framework_for_computing_smooth_strainlines} is
%evidently insufficient.
%
\input{mainmatter/results/figures/error_figures/xi2_error_fixed_steplength.tex}


\input{mainmatter/results/figures/error_figures/xi2_error_both.tex}

%\clearpage
