\subsection{Computed deviations in the flow maps}
\label{sub:computed_deviations_in_the_flow_maps}


Figure~\ref{fig:flowmap_err_fixed} indicates that the $\rmsd$ of the flow maps,
as defined in equation~\eqref{eq:rmsdflowmap} follow the expected power laws
of the numerical errors for the various singlestep integrators. This is
as expected, because the flow maps are found by direct application of the
numerical integrators. In addition, the flow map $\rmsd$ seem to exhibit
boundedness from below by the counteracting accumulated floating-point
arithmetic error, as hypothesized in
\cref{sub:on_the_choice_of_numerical_step_lengths_and_tolerance_levels}.
This effect is most prominent in the error curves of the Kutta and
classical Runge-Kutta schemes, which inflect upwards for sufficiently small
step lengths. Based on this evidence, the chosen step lengths appear to be
appropriate. Although neither the $\rmsd$ of the Euler scheme nor that of the
Heun scheme appears to have reached their respective inflection points, the
trends nevertheless indicate that there would not be much to gain in terms
of accuracy before the accumulated flating-point errors outweigh the inherent
precisions of the two schemes.

Figure~\ref{fig:flowmap_err_both} indicates that a peak in numerical accuracy
did not materialize for the adaptive stepsize methods, for the considered
numerical tolerance levels. The figure also reveals that the Bogacki-Shampine
3(2) scheme is the `cheapest' in terms of the required number of function
evaluations  when only a very crude approximation is needed, that is, using a
very high numerical tolerance level. Furthermore, the its higher order sibling
is able to keep up with the Dormand-Prince 8(7) method regarding efficiency,
until a mean of approximately $600$ function evaluations. This complies with the
notion that the Bogacki-Shampine 5(4) method is highly optimized and practically
behaves as an even higher order order method, as mentioned in
\cref{sub:the_runge_kutta_methods_under_consideration}.
Interestingly, there does not seem to be a direct correspondence between the
$\rmsd$ of the flow maps, and that of the computed LCS curves, presented in
\cref{sub:computed_deviations_in_the_lcs_curves}.

\input{mainmatter/results/figures/error_figures/flowmap_error_fixed_steplength.tex}

\input{mainmatter/results/figures/error_figures/flowmap_error_both.tex}
