\section{The LCS curves obtained using the different schemes}
\label{sec:the_lcs_curves_obtained_using_the_different_schemes}

Here, the repelling LCS curves found by means of the variety of
numerical integration schemes under consideration, are presented. All of the
LCS curves obtained for a given integrator, for all numerical time step lengths
or tolerance levels are included in one figure, where the reference LCS, shown
in figure~\ref{fig:referencelcs}, is also included in order to facilitate
visual comparison. The idea is that any LCS curve that deviates from the
reference, will reveal itself by not conforming perfectly. The LCS curves
resulting from the singlestep methods are presented in
\cref{fig:lcs_euler,fig:lcs_rk2,fig:lcs_rk3,fig:lcs_rk4}, whereas the ones
found by virtue of the embedded, that is, adaptive stepsize, methods, are
shown in~\cref{fig:lcs_rkbs32,fig:lcs_rkbs54,fig:lcs_rkdp54,fig:lcs_rkdp87}.

\subsection{LCS curves stemming from singlestep methods}
\label{sub:lcs_curves_stemming_from_singlestep_methods}



\input{mainmatter/results/figures/lcs_figures/euler.tex}
\input{mainmatter/results/figures/lcs_figures/rk2.tex}
\input{mainmatter/results/figures/lcs_figures/rk3.tex}
\input{mainmatter/results/figures/lcs_figures/rk4.tex}

The LCS curves obtained from the Euler scheme using relatively large step
lengths are the most obviously incorrect ones, in comparison to the reference.
There appears to be very little separating the performance of the other
singlestep methods. Interestingly, the Heun scheme appears to yield more
accurate results than the Kutta scheme for all the considered step lengths.
This is somewhat unexpected, because, as mentioned in
\cref{sub:the_runge_kutta_methods_under_consideration}, the Heun scheme is
\nth{2}-order accurate in the time step, whereas the Kutta scheme is
\nth{3}-order accurate. On a less surprising note, seeing as it is \nth{4}-order
accurate, the classical Runge-Kutta scheme produces very accurate curves for
all step lengths.

\subsection{LCS curves stemming from adaptive stepsize methods}
\label{sub:lcs_curves_stemming_from_adaptive_stepsize_methods}

Note that, although numerical tolerance levels of $10^{-1}$ through to
$10^{-10}$ were investigated, the figures presented here do not show LCS curves
for $\textnormal{tol}=10^{-1}$. This is due to the fact that, for all
embedded methods, none of the resulting domains $\mathcal{U}_{0}$ bore
a great amount of resemblance to the reference. This resulted in very disparate
behaviour of strainlines, to the extent that the resulting strain systems were
too dissimilar to the reference to warrant any meaningful comparisons. The
erroneous $\mathcal{U}_{0}$ domains are shown in~\cref{fig:u0_dom_errs}
(compare to~\cref{fig:u0_domain}).
%In fact, most likely because of the large density of points in these domains,
%keeping track of the neighbors for each individual strainline and its
%intersections with the lines in $\mathcal{L}$ turned out to be an unconquerable
%task. It is known that storing the necessary data requires in excess of 28 GB
%of memory, at which point even the NTNU supercomputer, Vilje, proved
%insufficient.

Notably, all of the embedded integration methods result in LCSs which are
dissimilar from the reference for the tolerance level $\textnormal{tol}=10^{-2}$,
where only the curve obtained by the Dormand-Prince 8(7) scheme is anything
alike the reference. More curiously, the Bogacki-Shampine 3(2) method appears
to outperform its theoretically more accurate sibling --- the
Bogacki-Shampine 5(4) method --- even for relatively
large tolerance levels. This is visible in the
vicinity of the `$\cup$' shape near $y=0.4$ in~\cref{fig:lcs_rkbs32,fig:lcs_rkbs54},
for instance. Furthermore, the Dormand-Prince 5(4) method appears to yield more
accurate LCS curves in general, than the Bogacki-Shampine 5(4) method. This
is somewhat surprising, given that the latter scheme is supposedly more accurate
than other methods of similar order, as mentioned in
\cref{sub:the_runge_kutta_methods_under_consideration}. Lastly, the
Dormand-Prince 8(7) method appears to produce very accurate results for all
tolerance levels smaller than $10^{-2}$, and, as previously mentioned, gives
the most correct LCS curve for the tolerance level $10^{-2}$
--- which is to be expected, as it is the highest order method among the ones
considered, after all.
\vspace{\fill}

\input{mainmatter/results/figures/domain_figures/all_erroneous_domains.tex}

\input{mainmatter/results/figures/lcs_figures/bs32.tex}
%\clearpage
\input{mainmatter/results/figures/lcs_figures/bs54.tex}
\input{mainmatter/results/figures/lcs_figures/dp54.tex}

\input{mainmatter/results/figures/lcs_figures/dp87.tex}


