\begin{framed}
    \begin{itemize}
        \item \sout{Approach does produce a consistent LCS picture for all numerical integrators considered, provided sensible
            integration steps or tolerance levels are chosen. This indicates that the calculation of this type
        of transport barrier is not particularly sensitive to the integration method.}
    \item \sout{Most crucial part: Compute advection correctly. The error of the computed strain eigenvalues scales
                like the advection error, while the error in the strain eigendirections is negligible for all time
                steps. The computed LCSs seem to exhibit sensitive dependence on the calculation of the eigenvalues,
            which is not particularly surprising, considering the LCS conditions of eq. 3.13 }
                \begin{itemize}
                    \item \sout{The precision of the strain eigendirections is likely a direct consequence of the
                        auxiliary tracers in their computation.}
                \end{itemize}
            \item \sout{The number of points in the $\mathcal{U}$ domain is different to the one obtained by
                Haller et  al, by approximately $3\%$. This, however, is likely related to how the
            grid of tracers was set up. In their paper, Farazmand and Haller do not provide details.}
        \item \sout{Notably, the reference LCS as shown in figure (.) consists of seven different strainline
                segments, unlike the structure found by Haller et al., which is claimed to be a single
            coherent strainline.}

        \item \sout{The time lenghts or tolerance levels should be chosen based on the system under consideration.
            In particular, an individual time step should not be so large that too much detail in the local,
        instantaneous velocity field is glossed over.}
    \item \sout{Possible approach in order to refine the computations: Use fewer `base' tracers, and an \emph{adaptive multigrid method}. Main con: May lead to inconsistent centered differencing when approaching
        the Jacobian etc. of the flow map}
        \item \sout{Regarding the completely different domains obtained via the embedded methods for the highest tolerance level:
            Check if this corresponds th the error in the flow map larger than some threshold. If (likely) so, this is
        further evidence that one should exhert great effort in computing the flow map correctly.}
    \item \sout{In terms of application to discrete velocity data sets, where both spatial and temporal interpolation
                    may be necessary, the interaction between interpolation and integration scheme is likely to
                    have a great overall impact. In particular, the underlying interpolation scheme sets a lower accuracy
                    bound for the entire advection process, practically enforcing restrictions on the numerical step
                    lengths or tolerance levels which can be considered sensible. This is also based on the considered
                velocity field.}
            \item \sout{Regarding the incompressibility of the velocity field}
                    \begin{itemize}
                        \item \sout{Incompressibility property is conserved until approximately $t=5$ units. Generally can't expect
                            this property to hold numerically over time.}
                        \item \sout{There is no assurance that neighboring tracers will remain nearby after the advection,
                                leaving the finite difference approximation of the local strain and stretch invalid. This
                                issue is expected to be most prominent in regions of high repulsion, which are precisely
                            where accuracy is imperative.}
                    \end{itemize}
        \item Parameter choices
            \begin{itemize}
                \item Numerical step lengths and tolerance levels
                \item The use of RK4 for all strainline iterations
                    \begin{itemize}
                        \item Alternative: Use a high order adaptive step method, paying close attention to the strainline curvature
                    \end{itemize}
                \item \sout{Represent tracer positions by means of higher precision floating-point numbers}
                        \begin{itemize}
                            \item \sout{Yields more accurate finite difference approximations}
                            \item \sout{Requires more memory, i.e., less tracers can be advected, leading to
                                        a more granular flow map representation. Perhaps most relevant for
                                        well-behaved velocity fields, for which the LCS approach is not really
                                        practically relevant --- why not simply advect a few particles and see where they
                                    end up?}
                        \end{itemize}
            \end{itemize}
        \item The identification process of local strain maximizing strainlines
            \begin{itemize}
                \item The cutting of strainline tails
                \item Lines in $\mathcal{L}$ $\rightarrow$ not necessarily a robust approach for general flows
                \item Alternative: Clustering algorithm
                \item Approach where both the strainline lengths and avg $\lambda_{2}$ are taken into consideration,
                    i.e., selecting a longer line with slighly smaller avg $\lambda_{2}$ over a shorter line with larger.
            \end{itemize}
        \item The special linear interpolation used for the eigenvectors
            \begin{itemize}
                \item Possible alternative: Higher order interpolation
            \end{itemize}
            \item Chose mean error, rather than max, because:
                \begin{itemize}
                    \item In terms of flow map: Error should be limited from above, by the domain extent, as the normal component
                        of the velocity at the boundaries is zero
                        \item Generally no way of telling where the maximum error occurs
                \end{itemize}
                \item Regarding the use of $\rmsd$:
                    \begin{itemize}
                        \item A practical application of empirical standard deviation
                            \item The distribution of errors across the domain is unknown. From the Central Limit thm., however,
                                the mean error is nearly distributed as a Gaussian.
                    \end{itemize}
    \end{itemize}
\end{framed}
