\section{Remarks on the computed LCSs}
\label{sec:general_remarks}

Overall, the use of a variational framework for computing LCSs appears to
produce robust and consistent LCS curves for all of the numerical integration
schemes considered here, subject to usage of a sufficiently small numerical
step length or tolerance level. This is apparent from the LCS curves shown
in~\cref{fig:lcs_euler,fig:lcs_rk2,fig:lcs_rk3,fig:lcs_rk4,fig:lcs_rkbs32,%
fig:lcs_rkbs54,fig:lcs_rkdp54,fig:lcs_rkdp87} conforming visually with the
reference LCS, and the computed errors in the various LCS curves --- shown
in~\cref{fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both} --- being small.
Furthermore, this indicates that calculations of this kind
are not particularly sensitive to integration method. Note that the
$\mathcal{U}_{0}$ domains obtained by means of the adaptive stepsize methods
for $\textnormal{tol}=10^{-1}$, shown in figure~\ref{fig:u0_dom_errs}, differ
greatly from the reference domain (see~\cref{fig:u0_domain}), which correlates
well with the computed $\rmsd$ of the flow maps, shown in
figure~\ref{fig:flowmap_err_both}. In particular, the error in the flow map for
the aforementioned tolerance level is of order $1$ for all of the embedded
methods. This error is comparable to the extent of the computational domain,
that is, $\mathcal{U}=[0,\hspace{0.5ex}2]\times[0,\hspace{0.5ex}1]$. Naturally,
this leads to a drastically different system.

The above observation implies that the most crucial part of the computation is
advecting the tracers accurately. As mentioned in
\cref{sub:computed_deviations_in_the_strain_eigenvalues_and_vectors}, the errors
in the computed strain eigenvalues and -vectors scale like the error in the
computed flow map. This is to be expected, as the eigenvalues and -vectors are
essentially found by applying finite differences to the flow map. The
eigenvalues and -vectors play a crucial in the LCSs identification process, as
is evident from the numerical reformulation of the LCS existence theorem, which
is given in~\cref{eq:numericalexistence}. The error in the computed strain
eigenvectors, is consistently two to three orders of magnitude smaller than the
error in the eigenvalues. This is most likely due to them being computed based
on the set of \emph{auxiliary} tracers, which per construction results in more
accurate finite differences.

From inspection of~\cref{fig:lcs_rmsd_fp_nn_fixed,fig:lcs_rmsd_fn_nn_fixed,%
fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both}, it becomes clear that the
error of the strainline components identified as LCS constituents, for some
configurations with small step lengths or tolerance levels, is dominated by
a seemingly constant contribution of the order $10^{-4}$. For instance,
for the Dormand-Prince 8(7) method,
\cref{fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both} show that for the three
lowest tolerance levels --- $\textnormal{tol}=10^{-8}$, $10^{-9}$ and
$10^{-10}$, respectively --- the $\rmsd$ is of order $10^{-4}$, whereas it is
considerably smaller --- of order $10^{-7}$ --- for $\textnormal{tol}=10^{-7}$.
This is surprising, as we expect the error to decrease when the tolerance level
is lowered.  Notably, this occurs for larger step lengths or tolerance levels,
respectively, than the corresponding turning points for the error in the flow
maps (i.e., the points at which the error in the flow maps increases with
decreasing step length or tolerance level). However, close inspection of
\cref{fig:lcs_rk2,fig:lcs_rk3,fig:lcs_rk4,fig:lcs_rkbs32,fig:lcs_rkbs54,%
fig:lcs_rkdp54,fig:lcs_rkdp87} reveals that, for the numerical step lengths
and tolerance levels which correspond to the same level of error in the LCS
curves (see~\cref{fig:lcs_rmsd_fp_nn_fixed,fig:lcs_rmsd_fn_nn_fixed,%
fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both}), the computed LCS
approximations are made up of seven different strainline segments each. This
is unlike the reference LCS, which, as previously mentioned, consists
of \emph{eight} strainline segments.
\clearpage
\Cref{fig:lcserroroscillations} shows the computed LCS approximations together
with the reference LCS, for the Dormand-Prince 8(7) method with tolerance
levels $\textnormal{tol}=10^{-5}$ through to $\textnormal{tol}=10^{-8}$ ---
the tolerance levels for which oscillations are visible in the computed errors
of the LCS curves (see~\cref{fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both})
--- for a subset of the computational domain $\mathcal{U}$. Notice in
particular that the second reference LCS curve segment (counting from the top
and downwards) is present in the approximations for $\textnormal{tol}=10^{-5}$
and $10^{-7}$, but \emph{not} for $\textnormal{tol}=10^{-6}$ and $10^{-8}$.
However, the two topmost reference LCS curve segments shown in the figure
are sufficiently close for this error not to be identified as a false negative.
Thus, when computing the LCS $\rmsd$, the offset between the LCS curve segment
corresponding to the topmost one shown in the figure, and the second topmost
reference LCS curve likely dominates the other contributions. Furthermore, the
fact that this offset is nearly constant explains the apparently identical LCS
$\rmsd$ in the cases where the computed LCS approximation consists of seven
strainline segments (as shown in
\cref{fig:lcs_rmsd_fp_nn_fixed,fig:lcs_rmsd_fn_nn_fixed,%
fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both}). For all of the numerical step
lengths and tolerance levels which result in the same level of LCS $\rmsd$, the
situation is the same as for the  tolerance levels $\textnormal{tol}=10^{-6}$
and $\textnormal{tol}=10^{-8}$, shown in~\cref{fig:lcserroroscillations}. Figures
showing the same level of  detail for the remaining integration methods have
thus been omitted for brevity. Note that the two reference LCS curve
segments in question nearly overlap, and their $\overline{\lambda}_{2}$ differ
by less than 5\%; thus, the absence of the second topmost reference LCS curve
segment would likely not have severe consequences for predictions regarding
the overall flow in the system.

\input{mainmatter/discussion/figures/lcserroroscillation.tex}

Following the above discussion, the results obtained here do not indicate
the existence of a lower sufficiency threshold in terms of the required
advection accuracy, beneath which the computed LCS curves do not become more
precise. Any such threshold would, however, likely only be valid for the LCS
curves of this particular velocity field --- that is, the system given by
\cref{eq:doublegyre,eq:doublegyrefuns,eq:doublegyreparams} --- which,
admittedly, appear quite robust overall. Other, more volatile systems, would
probably require more accurate advection. Further investigations
for a wider range of systems could result in valuable insight. Should such an
advection threshold exist, and be linked to the scales of the underlying system,
it would naturally be of great significance when investigating generic transport
systems by means of a similar variational LCS approach. However, there is
no apparent reason why such a threshold should exist.

%For the considered double gyre system, there appears to be a lower threshold
%in terms of the required advection accuracy, beneath which the computed LCS
%curves do not become more precise. This effect is apparent
%from inspection of
%\cref{fig:lcs_rmsd_fp_nn_fixed,fig:lcs_rmsd_fn_nn_fixed,%
%fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both}, where the error of strainline
%components identified as LCS constituents flattens abruptly. Notably, this
%occurs for larger numerical step lengths or tolerance levels, respectively,
%than the corresponding turning points for the error in the flow maps.
%For the double gyre system considered here, it appears that this advection
%accuracy threshold is of the order $10^{-6}$--$10^{-7}$, which follows from
%comparing~\cref{fig:lcs_rmsd_fp_nn_fixed,fig:lcs_rmsd_fn_nn_fixed,%
%fig:lcs_rmsd_fp_nn_both,fig:lcs_rmsd_fn_nn_both} to
%\cref{fig:flowmap_err_fixed,fig:flowmap_err_both}. In particular, for flow maps
%with $\rmsd$ of $10^{-7}$ or lower, the $\rmsd$ of the LCS curves appears to
%not decrease further as the flow map precision increases.
%
%However, because a similar flattening of the $\rmsd$ for the strain eigenvalues
%and eigenvectors is not apparent in
%\cref{fig:lmbd2_err_fixed,fig:lmbd2_err_both,fig:xi2_err_fixed,fig:xi2_err_both},
%one may infer that this threshold is likely only valid for the LCS curves
%of this particular velocity field --- that is, the system given by
%\cref{eq:doublegyre,eq:doublegyrefuns,eq:doublegyreparams} %
%%, for which LCSs are
%%found by means of the procedure described in
%%\cref{sec:advecting_a_set_of_initial_conditions,%
%%%    sec:calculating_the_cauchy_green_strain_tensor,%
%%%sec:identifying_lcs_candidates_numerically}
%--- which
%appear quite robust. The same flow map accuracy threshold probably does not
%suffice for other, more volatile flow systems. Investigating
%this further for a wider range of systems could result in valuable insight.
%Should such a threshold be valid in general, it would naturally be of great
%significance when investigating generic transport systems by means of a
%variational LCS approach. Admittedly, there is no apparent reason why
%this should be the case.
%
The double gyre model considered in this project is obviously not representative
of generic systems, in terms of the exact numerical step lengths or tolerance
levels necessary in order to obtain correct LCSs with a certain
degree of confidence. It does, however, indicate that these quantities should
be chosen based on the considered system. For a fixed stepsize integration
scheme, any single integration time step should not be so large that \emph{too}
much detail in the local and instantaneous velocity field is glossed over.
Similar logic applies when adaptive stepsize methods are used, although it
may be more difficult to enforce, depending on how the step length
update is implemented. One possibility in terms of choosing the time step, is to
find a characteristic velocity for the system, and choose the time step small
enough so that a tracer moving with the characteristic velocity never traverses
a distance greater than the grid spacing, when moving from one time level to
the next.
\clearpage
The computed reference LCS for the double gyre system considered here, shown in
\cref{fig:referencelcs}, is made up of \emph{eight} different strainline
segments. The LCS presented in the article by \textcite{farazmand2012computing}
is claimed to consist of a \emph{single} strainline segment. Comparing the two
curves visually, however, indicates that the resulting LCSs are similar.
Likewise, the domain $\mathcal{U}_{0}$, shown in figure~\ref{fig:u0_domain},
strongly resembles the one found by \citeauthor{farazmand2012computing}.
Nevertheless, the total number of points in the domain computed here is
approximately two percent larger than what \citeauthor{farazmand2012computing}
found. These discrepancies could originate from different conventions in terms
of generating the grid of tracers. Notably, \citeauthor{farazmand2012computing}
fail to provide a description of their approach.

When computing transport based on discrete data sets, such as snapshots of the
instantaneous velocity fields in oceanic currents, spatial and temporal
interpolation becomes necessary. Together with the inherent precision of the
model data, the choice of interpolation scheme(s) sets an upper bound
in terms of the accuracy with which tracers can be advected. For such cases,
the interaction between the integration and interpolation schemes could
be critical --- both in terms of computation time and memory requirements,
aside from the numerical precision. Independently of the scales at which
well-resolved LCS information is sought in this kind of system, the
aforementioned effects warrant further investigation.
