For the double gyre system considered here, the calculation of its LCSs
does not exhibit particularly sensitive dependence to
the choice of numerical integration method used for the initial tracer
advection. However, this could be a consequence of the LCSs present within the
system being robust under the chosen parameter values. This view is supported by
the fact that the errors in the computed LCS curves quickly flattened for
sufficiently small integration time steps or tolerance levels, while a similar
effect did not manifest itself for the computed flow maps or strain eigenvalues.
Based upon the errors in the computed flow maps, however, high order integration
methods are generally advisable for more generic flow systems --- they result in
more efficient calculations, which are less susceptible to numerical round-off
error. Moreover, the fact that the same strainline segments were identified as
LCSs even for quite large errors in the strain eigenvalues, suggests that the
numerical implementation of the variational principles in order to find LCSs is,
in itself, robust.

There is, however, room for further research with regards to the numerical
implementation of one of the LCS existence conditions (i.e., that of identifying
the locally most repelling strainlines), derived from their variational theory.
In particular, to our knowledge, a general, robust numerical implementation is
yet to be described in the literature. The approach used in this project
involves a set of parameters independent of the overall flow. The parameter
values were chosen based upon careful inspection of the system under
consideration, in order for the LCS curves obtained here to conform with those
found in the literature. A suggested alternative approach, which was not
investigated as part of this project, would be to utilize a sort of numerical
clustering algorithm rather than resorting to similar (to some extent)
subjective considerations to the ones which were employed here.

While no numerical integration scheme stood out as superior in general,
using higher order methods invariably resulted in more efficient calculations.
Accordingly, the use of higher order methods is generaly advisable. Equally
important as the choice of integration scheme, however, is the choice of
numerical step length or tolerance level. Ideally, this should be selected based
upon physical considerations of the system at hand. For instance, one could
identify a characteristic velocity, then tune the step length or tolerance level
such that no tracers moving with said velocity ever moves further than a
characteristic grid spacing, when moving from one time step to the next. This,
in order to ensure that the local instantaneous dynamics are resolved properly.
To what scale the microscopic behaviour should be resolved depends on the scale
to which detailed information regarding the flow is sought. Regarding most
real-life applications, this is also dictated by the the sampling frequency ---
or, more likely, model output --- of the discrete data samples (spatial and
temporal alike).

On that note, when investigating transport systems for which the available
data sets are discrete, the choice of numerical \emph{interpolation} scheme
will generally also impact the calculations of LCSs. This depencence has not
been investigated as part of this project. However, similar reservations
as for the integration time step or tolerance levels are also applicable to
the spatial and temporal sampling frequencies involved in the interpolation.
Furthermore, the choice of integration scheme should be made in relation
to the interpolation method. For instance, it likely makes little sense to use a
\nth{5}-order accurate integration scheme in tandem with a \nth{3}-order
accurate interpolator. Put simply, the interaction between interpolation
and integration schemes in the analysis of discrete data sets is of great
interest for practical applications of the LCS theory, and warrants
further investigation.
