\subsubsection{Suggestions for further work}

A plethora of numerical integration methods were not tested as a part of this
project. An entirely different class of ODE solvers to the ones considered here,
namely linear multistep methods, could have various desirable properties. In
particular, methods of this kind have memory. They take the velocities at
previous steps into account, which results in fewer velocity field evaluations
being necessary. If the velocity field is computationally expensive to evaluate,
the use of multistep methods opens up the possibility of using smaller numerical
step lengths than for singlestep methods without sacrificing computation time.
In short, multistep methods function quite differently to the memoryless
solvers considered here. Interestingly, both fixed and adaptive stepsize linear
multistep methods exist, and, by means of known recurrence relations, one may in
principle design multistep methods of arbitrary order
\parencite[chapter III]{hairer1993solving}. Research with regards to the
usefulness of such methods in the context of LCS detection --- in addition to
investigations on robust, numerical implementation of the as yet somewhat
ambiguous LCS existence condition regarding the identification of the locally
most repelling strainlines --- could reasonably be conducted based upon
the double gyre system examined in this project.
%Investigations regarding a robust, numerical implementation of one of the
%LCS existence conditions could also be conducted based upon the double gyre
%system considered here.

Curiously, in several quite recent research articles regarding the detection
of LCSs for real-world transport systems, high order adaptive step size
integration methods are used in conjunction with interpolated velocity fields,
seemingly with little awareness as to whether or not the use of high order
integration methods are warranted. For instance, the article by
\textcite{vansebille2018lagrangian} provides a quite thorough discussion of the
currently available computational tools for the advection of tracers, but lacks
a detailed treatment of numerical integration methods. Another example is
the article by \textcite{gough2017persistent}, where the authors state that they
used a high order adaptive step size Runge-Kutta method in combination with
a cubic interpolation routine in order to investigate oceanic transport
patterns in the northwestern gulf of Mexico; but fail to motivate their choices
of integration and interpolation methods. The results of this project indicate
that adaptive integration schemes of lower order (such as the Bogacki-Shampine
3(2) method) can be more efficient than higher order adaptive schemes when
somewhat crude approximations are sought (that is, using a relatively high
numerical tolerance level). In particular, velocity fields which are
interpolated by means of cubic splines are sufficiently smooth for integration
methods up to third order to work optimally.

Based on the above, investigations with regards to the interaction between
numerical integration and interpolation schemes in the context of LCS detection
are appealing, due to their innate relation to real-world applications.
For my own future work, I think that it would be natural to continue
the study of LCSs. In particular, I find the prospect of analyzing LCSs in
three-dimensional flow systems alluring. To my knowledge, the three-dimensional
formulation of the variational LCS approach has not yet been explored in the
literature. Thus, this could potentially accelerate the recommendability of
the use of LCS theory to describe systems which can not reasonably be regarded
as two-dimensional. Moreover, examining a broader range of systems seems like
a logical extension of the work I have conducted as part of this project.

%Furthermore, under the assumption that I am able to
%find a three-dimensional velocity field for which there exists a plane with a
%similar double gyre field to the one which has been investigated in detail in
%relation to this project, the results obtained here could prove useful
%with regards to verification.

%Moreover, there are a myriad of integration methods which were not investigated
%here. Aside from the many singlestep and embedded methods, there also exists
%an entirely different class of ODE solvers; namely, linear multistep methods.
%These methods have memory, and thus function quite differently from the
%ones considered for this project. Interestingly, both fixed and adaptive
%stepsize linear multistep methods exist, and, by means of known recurrence
%relations, one may in principle design multistep methods of arbitrary order
%\parencite[chapter III]{hairer1993solving}. A study of linear multistep
%integration methods can be viewed as a natural extension to the study of LCSs
%in three-dimensional flows, as suggested above.





%Regarding my own future work, I find the prospect of analyzing
%LCSs in three-dimensional flow systems alluring. To my knowledge, the
%three-dimensional formulation of the variational LCS approach has not yet
%been explored in detail in the literature. Thus, this could potentially
%accelerate the recommendability of the use of LCS theory to describe systems
%which can not reasonably be regarded as two-dimensional. Moreover, I think
%that it would be natural to continue working on LCSs, considering that I have
%developed a good understanding of the underlying physical principles through the
%work I have conducted so far --- not to mention the fact that I have become
%quite adept regarding the use of supercomputers for scientific purposes, which
%I most certainly expect will become necessary, in view of the added inherent
%mathematical complexity that follows by moving from two to three spatial
%dimensions.
